### 2011-02-21 trunk/deliverables/D4_3/fast_d4_3_2.tex
Ismael review: I fixed some mistakes, and make a few comments for improvement
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1277 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass{article}

\usepackage{fast_latex}
% yay! We can use any kind of funky diacritic:
\usepackage[utf8]{inputenc}

@@

\newcommand\deliverableNumber{D4.3.3}
\newcommand\deliverableTitle{Mechanisms for Gadget-Service Connections and Gadget Functionality}
\newcommand\deliverableTitleShort{Mechanisms for Gadget-Service Connections}
\newcommand\workpackageNumber{4}
\newcommand\workpackageTitle{Visual composition of screen-flow resources and interoperability with back-end Web Services}
\newcommand\authorOne{Ismael Rivera, NUIG}
\newcommand\authorTwo{Albert Zündorf, UniKas}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

@@

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		{\large \sffamily \authorTwo}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 26/02/2010}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
							\textcolor{white}{\textbf{Change description}}\\ \hline
%0.1 & 15.12.2008 & Knud M\"oller (NUIG) & template modelled \\ \hline
%0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
1.0 & 26.02.2011 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
 & & Albert Zündorf (UniKas) & \\ \hline
\end{tabular}
\end{small}

@@

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.3 is...}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ******} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web services, WSDL, REST, SOAP, WSMO}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

This deliverable is structured as follows. Section ~\ref{sec:introduction} states the goal, scope and structure of the document. Section ~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how these wrappers are build and then discovered to be reused in any gadget. Finally, Section ~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.

@@

Therefore, a complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds). These services and data sources need to be modelled and encapsulated within the platform in order to allow the discovery and use of them by other components (i.e. forms and operators). This is called Resource Adapter in the FAST platform. Figure ~\ref{fig:complex_gadget_architecture} illustrates their position and interaction with the rest of the building blocks within a gadget.

@@

To permit a high number of REST-based web services to be integrated to the FAST platform, the approach taken is from a manual development perspective. There is no need of a formal document such as WADL defining the web service, for this reason, building service wrappers for these services involve the correct understanding of the service by a human being and a tool to facilitate this task. This tool is explained in detail in Section ~\ref{sec:restful_web_services_wrapper_tool}.

@@

From a specific WSDL definition, a set of methods or operations can be easily extracted which encloses a set of inputs and outputs. These operations would be transformed into actions, and the inputs and outputs would be used to define Resource Adapter pre/postconditions. With the use of this information along with the XML Schema defining the types, a Resource Adapter can be semantically defined.

To integrate WSDL defined services into the FAST platform, we again propose an interactivce approach: We extend the tool for REST service wrapping with the ability of loading WSDL definitions of a service and this WSDL definition is then exploited for guidance of the user through the Resource Adapter construction, cf. Section ~\ref{sec:restful_web_services_wrapper_tool}.

@@

A web service in WSMO defines one and only one capability. The capability of a web service defines its functionality in terms of pre/postconditions, assumptions and effects. A web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, shared variables, precondition, postcondition, assumption, and effect.

Basically, the web service will offers to a client a postcondition when some conditions are met in the information space (precondition). Effects and assumptions can stay out of this study, since they will not be taken into account by any of the building blocks of a gadget.

Like WSDL defined services, we incorporate WSMO services into the FAST platform in an interactive storyboard driven approach. The FAST tool described in Section ~\ref{sec:restful_web_services_wrapper_tool} reads the WSMO specification of a service and exploits this information again for guidance for the contstruction of an Resource Adapter.

@@

In the building phase, every Resource Adapter is created sharing a common structure for screen components, such as forms, operators and resources. Hence, it will have a set of actions which will contain a set of preconditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their postconditions.

For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the catalog, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions. In a first step, the system will suggest screen components which have any of the preconditions which can be connected to any of the screen preconditions (the connection will be represented by an internal pipe), and the same would happen for the postconditions. Once a screen component is inserted into the screen, the system will take its definition into consideration to recommend new screen components which can be connected to the ones the screen is composed of.

In these bases, the mechanism is relatively similar to the screen recommender, used to find the best screens to create a screen flow. Once more, it is based on the pre/postconditions of a set of building blocks.

@@

\clearpage
\section{Web Service Wrapper Tool} % (fold)
\label{sec:restful_web_services_wrapper_tool}

In FAST, the wrapper tool will be in charge of the Resource Adapters' building phase. The building of Resource Adapters is done in three steps: in a first step, we define which pre- and postconditions are used to pass parameters between FAST building blocks and the Resource adapter. The second step is in charge of the construction of a service request and the third step analyses the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.

\subsection{Defining pre- and postconditions} % (fold)
\label{sub:resource_adapter_parameters}

We will demonstrate the use of the FAST Service Wrapper Tool on the example of a simple Formula1 service. In the first tab of the FAST Service Wrapper Tool, the user defines the name and the pre- and postconditions of the desired Resource Adapter, cf. Fig. \ref{fig:construct_pre_post_conditions}. For the precondition, we provide a name, a type, and an example value. The type of a pre- or postcondition is chosen from a drop down list of available concepts. This list of available concepts is retrieved from the FAST ontology catalog. The example value is used for test runs during subsequent construction steps. We keep track of example values provided for different domain concepts and propose them, when the domain concept is re-used e.g. in another Resource Adapter.

\begin{figure}
  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPrePostConds.png}
    \caption{Configuring pre- and postconditions of a Resource Adapter}
    \label{fig:construct_pre_post_conditions}
  \end{center}
\end{figure}

@@

In our simple case, the service request consists just of a URL that is easily retrieved from API documentation page of our Formula1 service. This URL consists of the service address \url{http:// ergast.com/api} followed by the year we are querying for followed by the category \url{drivers} we are interested in. The user may grab such an URL from the API documentation page of the targeted web service and drop it into the Request tab of the FAST Service Wrapper Tool, cf. Figure ~\ref{fig:construct_service_request}.  The FAST Service Wrapper Tool will than analyze the dropped URL and split it into its parts and provide a convenient form allowing to edit e.g. all the parameters of the URL. In our example the URL has no parameters so the parameter form shows only the base URL.

For WSDL or WSMO services, the FAST Service Wrapper Tool provides additional support for the construction of the request. The user may point the FAST Service Wrapper Tool to the WSDL or WSMO definition of the targeted service. The FAST Service Wrapper Tool then reads this definition and analysis it. With the help of this information, the FAST Service Wrapper Tool provides interactive guidance for selecting the desired operation from a service and then the FAST Service Wrapper Tool retrieves the operation descriptions and fills the request parameter form, adequately.

Now we have to connect the request parameters to the parameter values passed via the preconditions to the Resource Adapter. In our case, we replace the year part of the URL with a placeholder for the precondition season \url{<season>}, see Request URL input field at the top of Figure ~\ref{fig:construct_service_request}.

In our example, we use a simple GET based web service. The FAST Service Wrapper tool also supports the other kinds of URL operations as POST, PUT, etc. For POST requests, one usually passes parameters within a special post body. For unstructured POST bodies, we provide a text area allowing the user to fill in placeholders for references to precondition values. In case of structured POST bodies as in the case of WSDL and WSMO services the POST body uses a structured text format based e.g. on JSON or XML. In that case, the body text is again passed and the parameters are included in the request parameter form part. Then the user may easily use the request parameter form part to fill in precondition placeholders.

Figure ~\ref{fig:construct_service_request} shows the screen to construct the service requests.

@@

Below the request paramter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed request by sending it to the specified service address (via a server relay). Then the placeholders for preconditions are replaced with example values and the resulting request is shown below the parameter form. In addition, the request is send and the response is shown on the bottom of that page. This gives the user a fast feedback whether the constructed request works as desired.

@@

It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types. However, every building block accepts any (complex) concept as a pre/postcondition, so the wrapper tool needs to be adapted to permit any concept as input or precondition, extracting the required text to construct the request. This should be easy to achieve before the project review.

@@

Once the service request is constructed and sent to the service provider, it will send back a response. This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON among others. To continue the example started in the previous section, the response of the Formula1 service will is in XML. However, the FAST wrapper tool supports JSON, similarly.

@@

Figure ~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.

\begin{figure}
  \begin{center}
      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRulesF1.png}
    \caption{Interactive, rule based transforming of an XML response to FAST facts (photoshoped screen dump)}
    \label{fig:response_service_execution}
  \end{center}
\end{figure}

Once the service response, in XML or JSON format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}. To construct this interactive object tree, the XML or JSON document is parsed into a DOM, and a simplified tree representation of that DOM is built up. This tree representation of the XML/JSON data is used as an input to construct transformation rules.

A transformation rule is used to analyse the XML/JSON data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks. A transformation rule is composed of three elements, cf the middle part of Figure ~\ref{fig:response_service_execution}. First, the \textit{from} field indicates the XML elements to be translated by rule. These XML elements are identified by the tagname of a DOM element from the XML document. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be trigger from the transformation rules is:
\begin{description}
	\item[action \emph{createObject}] specifies the creation of a new fact object. The type of that new fact is provided in the third compartment. In the example being explained, the root rule searches for XML elements with tagname \emph{DriverTable} and for each such element a \emph{List} fact is created. The resulting facts are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
	\item[action \emph{fillAttributes}] does not create a new fact but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tagname \emph{FamilyName}. Note, the rule is a sub-rule of the second rule, which generates \emph{Person} facts. Thus, the sub-rule searches for \emph{FamilyName} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Driver} rule may just have been applied to the first \emph{Driver} element of the XML data. Then, the \emph{FamilyName} rule is applied only to the first \emph{Driver} sub-tree of the XML data and thus it will find only one \emph{FamilyName} element in that sub-tree. The value of that \emph{FamilyName} element is then transfered to the \emph{name} attribute of the corresponding \emph{Person} fact. Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g. to \textit{words} 1 through 3. It is also possible to combine constant text and elements of multiple XML tree elements. A \textit{from} field may also refer to multiple tags of the source tree, e.g. to combine a full name from a first name and a family name field.  
	\item[action \emph{dummy}] does not create or modify any facts but such rules are just used to narrow the search space for their sub-rules. For example, in the Amazon case, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
\end{description}

Since FAST is storyboard oriented, the FAST Service Wrapper Tool follows the storyboard idea as well. Any time, a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, FAST is semantic-driven, therefore, the service wrapper designer retrieves the domain-specific types from the FAST ontology catalog server together with the structure of each type, i.e. together with a description of the attributes of each fact. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the fact types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the fact type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the facts generated by the transformation rules conform to the fact types defined in the corresponding FAST ontology. This helps to ensure that the facts generated by the designed Resource Adapter will be compatible for preconditions of subsequent filter steps and or screens.

@@

Once the wrapping of a service has been defined and tested in the FAST Service Wrapper Tool, it generates an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget. In addition, a description of the new building block is added to the FAST catalog. Thereby, the new Resource Adapter becomes available for the construction of new screens and screen flows.

@@

There is a number of related work that we have consider to use for service wrapping instead of building a Service Wrapper
Tool ourself. However, due to the special needs of the Gadget Development Process it turned out that we were not able 
to use an existing approach directly but we were only able to grab 
ideas from existing approaches and come up with our own implementation. 
As discussed, service wrapping requires three steps. First we have to define the connection between the desired service wrapper and the other building blocks of FAST. This is achieved by the definition of pre- and postconditions in our tool. Since this is very FAST specific, finding an existing tool for this purpose did not work. However, this is a simple step anyhow requiring only 
limited implementation efforts.

>>>>

\documentclass{fast_latex}

% yay! We can use any kind of funky diacritic:
\usepackage[utf8]{inputenc}

@@

\newcommand\deliverableNumber{D4.3.2}
\newcommand\deliverableTitle{Mechanisms for Gadget-Service Connections and Gadget Functionality}
\newcommand\deliverableTitleShort{Mechanisms for Gadget-Service Connections}
\newcommand\workpackageNumber{4}
\newcommand\workpackageTitle{Visual composition of screen-flow resources and interoperability with back-end Web Services}
\newcommand\authorOne{Ismael Rivera (NUIG)}
\newcommand\authorTwo{Albert Zündorf (UniKas)}
\newcommand\authorThree{Knud M\"oller (NUIG)}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

@@

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		{\large \sffamily \authorTwo}\\
		{\large \sffamily \authorThree}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 27/02/2010}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
							\textcolor{white}{\textbf{Change description}}\\ \hline
%0.1 & 15.12.2008 & Knud M\"oller (NUIG) & template modelled \\ \hline
%0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
1.0 & 26.02.2010 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
 & & Albert Zündorf (UniKas) & \\ \hline
2.0 & 27.02.2011 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
 & & Albert Zündorf (UniKas) & \\
 & & Knud M\"oller (NUIG) & \\ \hline
\end{tabular}
\end{small}

@@

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.2 is...}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne, \authorTwo, \authorThree}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ?????} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web services, WSDL, REST, SOAP, WSMO}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

This deliverable is structured as follows. Section~\ref{sec:introduction} states the goal, scope and structure of the document. Section~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how these wrappers are build and then discovered to be reused in any gadget. Finally, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.

% subsection structure_of_the_document (end)

\subsection{Changes from Previous Version} % (fold)
\label{sub:changes_from_previous_version}

\textcolor{red}{To be done...}

% subsection changes_from_previous_version (end)

@@

Therefore, a complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds). These services and data sources need to be modelled and encapsulated within the platform in order to allow the discovery and use of them by other components (i.e. forms and operators). This is called Resource Adapter in the FAST platform. Figure~\ref{fig:complex_gadget_architecture} illustrates their position and interaction with the rest of the building blocks within a gadget.

@@

To permit a high number of REST-based web services to be integrated to the FAST platform, the approach taken is from a manual development perspective. There is no need of a formal document such as WADL defining the web service, for this reason, building service wrappers for these services involve the correct understanding of the service by a human being and a tool to facilitate this task. This tool is explained in detail in Section~\ref{sec:restful_web_services_wrapper_tool}.

@@

From a specific WSDL definition, a set of methods or operations can be easily extracted which encloses a set of inputs and outputs. These operations would be transformed into actions, and the inputs and outputs would be used to define Resource Adapter pre-/post-conditions. With the use of this information along with the XML Schema defining the types, a Resource Adapter can be semantically defined.

To integrate WSDL defined services into the FAST platform, we again propose an interactivce approach: We extend the tool for REST service wrapping with the ability of loading WSDL definitions of a service and this WSDL definition is then exploited for guidance of the user through the Resource Adapter construction, cf. Section~\ref{sec:restful_web_services_wrapper_tool}.

@@

A web service in WSMO defines one and only one capability. The capability of a web service defines its functionality in terms of pre-/post-conditions, assumptions and effects. A web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, shared variables, pre-condition, post-condition, assumption, and effect.

Basically, the web service will offers to a client a post-condition when some conditions are met in the information space (pre-condition). Effects and assumptions can stay out of this study, since they will not be taken into account by any of the building blocks of a gadget.

Like WSDL defined services, we incorporate WSMO services into the FAST platform in an interactive storyboard driven approach. The FAST tool described in Section~\ref{sec:restful_web_services_wrapper_tool} reads the WSMO specification of a service and exploits this information again for guidance for the construction of a Resource Adapter.

@@

In the building phase, every Resource Adapter is created sharing a common structure for screen components, such as forms, operators and resources. Hence, it will have a set of actions which will contain a set of pre-conditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their post-conditions.

For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the Catalogue, establishing which the screen's pre-/post-conditions are, and then the system will suggest several screen components which may satisfy these pre-/post-conditions. In a first step, the system will suggest screen components which have any of the pre-conditions which can be connected to any of the screen pre-conditions (the connection will be represented by an internal pipe), and the same would happen for the post-conditions. Once a screen component is inserted into the screen, the system will take its definition (screen components, pipes, etc.) into consideration to recommend new screen components which can be connected to the ones the screen is composed of.

For a further explanation on how the recommendation algorithms are implemented, please see~\cite{rivera2011catalogue_user_manual}.

@@

\clearpage
\section{Service Wrapper Tool} % (fold)
\label{sec:restful_web_services_wrapper_tool}

In FAST, the Service Wrapper Tool will be in charge of the Resource Adapters' building phase. The building of resource adapters is done in three steps: in a first step, we define which pre- and post-conditions are used to pass parameters between FAST building blocks and the resource adapter. The second step is in charge of the construction of a service request and the third step analyses the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific classes or concepts from the ontologies used within FAST by the building blocks' conditions.

\subsection{Defining pre- and post-conditions} % (fold)
\label{sub:resource_adapter_parameters}

We will demonstrate the use of the Service Wrapper Tool on the example of a simple \emph{Formula1} service. In the first tab of the Service Wrapper Tool, the user defines the name and the pre- and post-conditions of the desired Resource Adapter, cf. Fig. \ref{fig:construct_pre_post_conditions}. For the pre-condition, we provide a name, a type, and an example value. The type of a pre- or post-condition is chosen from a drop down list of available concepts. The list of available concepts, together with its properties and some example data (instances) is retrieved from the Catalogue. The example value is used for test requests during subsequent construction steps. We keep track of example values provided for different domain concepts and propose them, when the domain concept is re-used e.g. in another Resource Adapter.

\begin{figure}
  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPrePostConds.png}
    \caption{Configuring pre- and post-conditions of a Resource Adapter}
    \label{fig:construct_pre_post_conditions}
  \end{center}
\end{figure}

@@

In our simple case, the service request consists just of a URL that is easily retrieved from API documentation page of our Formula1 service. This URL consists of the service address \url{http://ergast.com/api} followed by the year we are querying for followed by the category \verb|drivers| we are interested in. The user may grab such an URL from the API documentation page of the targeted web service and drop it into the Request tab of the Service Wrapper Tool, cf. Figure~\ref{fig:construct_service_request}. The Service Wrapper Tool will than analyze the dropped URL and split it into its parts and provide a convenient form allowing to edit e.g. all the parameters of the URL. In our example the URL has no parameters so the parameter form shows only the base URL.

For WSDL or WSMO services, the Service Wrapper Tool provides additional support for the construction of the request. The user may point to the WSDL or WSMO definition of the targeted service. The Service Wrapper Tool then reads this definition and analyses it. With the help of this information, the Service Wrapper Tool provides interactive guidance for selecting the desired operation from a service and then it retrieves the operation descriptions and fills the request parameter form adequately.

Now we have to connect the request parameters to the parameter values passed via the pre-conditions to the Resource Adapter. In our case, we replace the year part of the URL with a placeholder for the pre-condition season \verb|<season>|, see Request URL input field at the top of Figure~\ref{fig:construct_service_request}.

In our example, we use a simple HTTP GET request. The Service Wrapper Tool also supports the other kinds of HTTP methods as POST, PUT or DELETE. For POST requests, one usually passes parameters within a special post body. For unstructured POST bodies, we provide a text area allowing the user to fill in placeholders for references to pre-condition values. In case of structured POST bodies as in the case of WSDL and WSMO services the POST body uses a structured text format based e.g. on JSON or XML. In that case, the body text is again passed and the parameters are included in the request parameter form part. Then the user may easily use the request parameter form part to fill in pre-condition placeholders.

\textcolor{red}{WSDL is usually used with SOAP and XML, JSON wouldn't be common in this case. And from my understanding, the user has to hand-write the POST body? therefore the SOAP message? that's doable, but pretty hard and tedious. - Ismael}

\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael}

Figure~\ref{fig:construct_service_request} shows the screen to construct the service requests.

@@

Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed request by sending it to the specified service address (via a server relay). Then the placeholders for pre-conditions are replaced with example values and the resulting request is shown below the parameter form. In addition, the request is sent and the response is shown on the bottom of that page. This gives the user a fast feedback whether the constructed request works as desired.

@@

It is worth pointing out that currently the Service Wrapper Tool is able to construct input ports for the wrappers using just basic types. However, every building block accepts any (complex) concept as a pre-/post-condition, so the wrapper tool needs to be adapted to permit any concept as input or pre-condition, extracting the required text to construct the request. This should be easy to achieve before the project review.

\textcolor{red}{I guess this paragraph will be rewritten...because my understanding is that we do support constructing wrappers with ``complex'' pre-/post-conditions and not just basic types. - Ismael}

@@

Once the service request is constructed and sent to the service provider, it will send back a response. This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON. To continue the example started in the previous section, the response of the Formula1 service will is formatted in XML. Th Service Wrapper Tool supports both JSON and XML.

@@

Figure~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.

\begin{figure}
  \begin{center}
      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRulesF1.png}
    \caption{Interactive, rule based transforming of an XML response to FAST facts}
    \label{fig:response_service_execution}
  \end{center}
\end{figure}

Once the service response has been retrieved, it is shown in the transformation tab as an interactive object tree, as it can be seeing on the left side of Figure~\ref{fig:response_service_execution}. To construct this interactive object tree, the XML or JSON document is parsed into a DOM, and a simplified tree representation of that DOM is built up. This tree representation of the XML/JSON data is used as an input to construct transformation rules.

A transformation rule is used to analyse the XML/JSON data and to generate domain-specific facts from concepts from the ontologies used by the pre-/post-conditions of the different building blocks. A transformation rule is composed of three elements, as the middle part of Figure~\ref{fig:response_service_execution} shows. First, the \textit{from} field indicates the XML elements to be translated by rule. These XML elements are identified by the tagname of a DOM element from the XML document. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be triggered from the transformation rules is:
\begin{description}
	\item[action \emph{createObject}] specifies the creation of a new fact object. The type of that new fact is provided in the third compartment. In the example being explained, the root rule searches for XML elements with tagname \emph{DriverTable} and for each such element a \emph{List} fact is created. The resulting facts are shown in a facts tree in the right of Figure~\ref{fig:response_service_execution}.
	\item[action \emph{fillAttributes}] does not create a new fact but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tagname \emph{FamilyName}. Note, the rule is a sub-rule of the second rule, which generates \emph{Person} facts. Thus, the sub-rule searches for \emph{FamilyName} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Driver} rule may just have been applied to the first \emph{Driver} element of the XML data. Then, the \emph{FamilyName} rule is applied only to the first \emph{Driver} sub-tree of the XML data and thus it will find only one \emph{FamilyName} element in that sub-tree. The value of that \emph{FamilyName} element is then transfered to the \emph{name} attribute of the corresponding \emph{Person} fact. Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g. to \textit{words} 1 through 3. It is also possible to combine constant text and elements of multiple XML tree elements. A \textit{from} field may also refer to multiple tags of the source tree, e.g. to combine a full name from a first name and a family name field.  
	\item[action \emph{dummy}] does not create or modify any facts but such rules are just used to narrow the search space for their sub-rules. For example, in the Amazon case, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
\end{description}

Since FAST is storyboard oriented, the Service Wrapper Tool follows the storyboard idea as well. Any time, a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, FAST is semantic-driven, therefore, the service wrapper designer retrieves the domain-specific types from the Catalogue server together with the structure of each type, i.e. together with a description of the attributes of each fact. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the fact types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the fact type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the facts generated by the transformation rules conform to the fact types defined in the corresponding FAST ontology. This helps to ensure that the facts generated by the designed Resource Adapter will be compatible for pre-conditions of subsequent filter steps and or screens.

\textcolor{red}{We may provide some analysis tool? is it implemented? otherwise I don't see it's worth to point this out, or maybe just in a ``future work'' section - Ismael}

@@

Once the wrapping of a service has been defined and tested in the Service Wrapper Tool, it generates an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget. In addition, a description of the new building block is added to the FAST Catalogue. Thereby, the new Resource Adapter becomes available for the construction of new screens and screen-flows.

\textcolor{red}{XML, HTML and Javascript? I don't get why XML and HTML is generated. The service wrapper should be just a javascript encapsulation of the web service, isn't it? HTML files should be for forms and XML for specific platform configuration files for deployment? - Ismael}

@@

There is a number of related work that we have consider to use for service wrapping instead of building a Service Wrapper
Tool ourself. However, due to the special needs of the Gadget Development Process it turned out that we were not able 
to use an existing approach directly but we were only able to grab ideas from existing approaches and come up with our own implementation. 
As discussed, service wrapping requires three steps. First we have to define the connection between the desired service wrapper and the other building blocks of FAST. This is achieved by the definition of pre- and post-conditions in our tool. Since this is very FAST specific, finding an existing tool for this purpose did not work. However, this is a simple step anyhow requiring only limited implementation efforts.
